% ---------------------------------------------------------------------------
% One-warehouse multi-retailer (OWMR) section.
%
% INPUT-ABLE FRAGMENT: this file is \input{} into the manuscript
% learning_inventory_control_policies_es.tex and reuses ITS preamble macros
% and TikZ styles (envbox, envtitle, envsupply/envstock/envsink/envlost,
% reqflow/matflow/demandflow/lostflow, ltlabel/reqlabel, costH/costP, the boxed
% "Per-period event sequence" panel, booktabs, \ding). It defines NO new styles
% and contains NO \documentclass.
%
% ALGORITHMIC / DATA-PROVENANCE NOTE (every number is sourced from the repo):
%   Instance + dynamics : src/problems/one_warehouse_multi_retailer/env.rs
%       (step_state: warehouse receipt -> allocation -> retailer receipt ->
%       demand -> partial backorder/emergency; warehouse holding charged
%       POST-emergency per Kaynov Eq.6), allocation.rs (proportional floor per
%       Kaynov Eq.8; min_shortage = max-min resulting position), references.rs
%       (TABLE_A3_INSTANCES; protocol 100 periods, 1000 reps, L_w=2, emergency
%       prob 0.8, h_w=0.5, h_r=1). The asymmetric/high-CV partial-backorder
%       instances kaynov2024_instance_{12,13,14} are studied here.
%   Comparator gate     : strongest in-repo grid-searched echelon base-stock +
%       allocation, re-scored on the held-out CRN block. The published Kaynov
%       PPO row (= -reward) is a REFERENCE learned comparator, NOT the gate.
%   Results             : outputs/one_warehouse_multi_retailer/asymmetric_learned/
%       *.json + scripts/one_warehouse_multi_retailer/ASYMMETRIC_LEARNED_RESULTS.md
%       (committed c976af5). instance_12 learned 1154.09 vs gate 1169.59
%       (+1.33%, paired +15.50 +/- 0.97, ~16 SEM); instance_13 learned 85974.79
%       vs gate 91890.25 (+6.44%, paired +5915.47 +/- 49.50, ~120 SEM);
%       instance_14 ties (trained xbest > gate -> deployed gate-reproducing
%       anchor). Published PPO 1118.92 / 79727.39 / 42835.02.
%   HONEST FRAMING      : learned beats the strongest IN-REPO HEURISTIC gate
%       beyond SEM on 2/3 and ties on the hardest; it does NOT beat the
%       published PPO (PPO remains strongest learned policy on every row).
% ---------------------------------------------------------------------------
\section{One warehouse, multiple retailers}
\label{sec:owmr}

\subsection{Problem formulation}
\label{sec:owmr-formulation}

Figure~\ref{fig:env-owmr} depicts this environment as a flow network --- the warehouse order and
its lead-time receipt, the downstream allocation of scarce warehouse stock to the retailers, each
retailer's lead-time receipt and demand draw, and where holding and the partial-backorder penalty
are charged.

\begin{figure}[t]
  \centering
  \resizebox{\textwidth}{!}{%
  \begin{tikzpicture}[x=1cm,y=1cm]
    % Two-echelon divergent system: one external supplier feeds a single
    % warehouse, which allocates scarce on-hand stock across K retailers (three
    % drawn). Order requests UP (dashed control); lead-time material DOWN (solid).
    % Each retailer faces its own demand; under the partial-backorder regime a
    % fraction of a shortage is backordered and the rest is lost.
    \node[envsupply] (sup)  at (0,0)      {external\\supplier};
    \node[envstock]  (wh)   at (6.2,0)    {warehouse\\IP $y^w_t$\\(lead $L_w$)};
    \node[envstock]  (r1)   at (13.0,3.0) {retailer $1$\\IP $y^1_t$};
    \node[envstock]  (r2)   at (13.0,0.0) {retailer $k$\\IP $y^k_t$};
    \node[envstock]  (r3)   at (13.0,-3.0){retailer $K$\\IP $y^K_t$};
    \node[envsink]   (d2)   at (19.4,0.0) {demand $D^k_t$};
    \node[envlost]   (b2)   at (19.4,-3.4){shortage:\\backorder $\beta$\\/ \emph{lost} $1-\beta$};

    % warehouse replenishment: dashed request UP, solid lead-time arrival DOWN
    \draw[reqflow] (wh.north west) to[bend right=30] (sup.north east);
    \node[reqlabel] at (2.9,1.9) {\ding{183}\, warehouse order};
    \draw[matflow] (sup.south east) to[bend right=10] (wh.west);
    \node[ltlabel]  at (3.0,-1.5) {\ding{185}\, receipt (placed $L_w$ ago)};

    % allocation warehouse -> retailers (solid material, scarce -> rationed)
    \draw[matflow] (wh.east) to[bend left=10]  (r1.west);
    \draw[matflow] (wh.east) -- (r2.west);
    \draw[matflow] (wh.east) to[bend right=10] (r3.west);
    \node[ltlabel] at (9.6,1.55) {\ding{186}\, allocate scarce\\on-hand (proportional\\/ min-shortage)};

    % one representative retailer's demand + partial backorder
    \draw[demandflow] (r2) -- (d2)
          node[midway,above=3pt,ltlabel] {\ding{187}\, demand};
    \draw[lostflow] (d2) -- (b2);

    \node[costH,above=8mm of wh.north] {holding $h_w(y^w_t)^+$};
    \node[costH,above=3pt of r1.north] {holding $h_r$};
    \node[costP,right=2pt of b2.east] {penalty $p\,b^k_t$};

    \begin{scope}[on background layer]
      \node[envbox,fit=(current bounding box)] (box) {};
    \end{scope}
    \node[envtitle,above=3pt of box.north,anchor=south]
          {One-warehouse, $K$-retailer environment (partial backorder, warehouse lead $L_w$)};
  \end{tikzpicture}%
  }
  \par\vspace{6pt}\centerline{%
    \tikz\node[draw=mydarkblue!55,rounded corners=4pt,fill=mydarkblue!4,inner sep=7pt]{%
      \begin{minipage}{\dimexpr\linewidth-18pt\relax}\centering
        {\small\bfseries\color{mydarkblue!85}Per-period event sequence}\\[4pt]
        {\small\begin{tabular}{@{}p{4.3cm}p{4.3cm}p{4.3cm}@{}}
          \ding{182}\,observe warehouse + retailer inventory positions & \ding{183}\,set warehouse order-up-to and per-retailer targets & \ding{184}\,warehouse receipt (placed $L_w$ ago) arrives\\[4pt]
          \ding{185}\,\emph{allocate} scarce warehouse on-hand across retailers (proportional / min-shortage) & \ding{186}\,retailer receipts arrive; demand $D^k_t$ drawn; a fraction $\beta$ of each shortage is backordered, the rest lost & \ding{187}\,charge $h_w$ on warehouse on-hand, $h_r$ on retailer on-hand, penalty $p$ on backorders\\
        \end{tabular}}
      \end{minipage}};}\par
  \caption{One-warehouse multi-retailer environment of \citet{kaynov2024owmr}, drawn as an
  event-based interaction. A single warehouse \emph{orders} from an external supplier (thin dashed
  control arrow) and receives the goods after a lead time $L_w$ (solid material arrow). Each period
  the warehouse \emph{allocates} its scarce on-hand stock across the $K$ retailers under a fixed
  rationing rule --- \emph{proportional} (floor rationing, Kaynov Eq.~8) or \emph{min-shortage} ---
  and each retailer receives its allocation after its own lead time, then faces demand $D^k_t$. Under
  the \emph{partial-backorder} regime a fraction $\beta$ of any retailer shortage is backordered and
  the remainder is lost. Warehouse holding $h_w$ is charged on warehouse on-hand stock (after
  emergency fulfilment, Kaynov Eq.~6), retailer holding $h_r$ on retailer on-hand stock, and a penalty
  $p$ on backordered units; the boxed panel gives the per-period event sequence
  (Section~\ref{sec:owmr-formulation}, Eq.~\eqref{eq:owmr-cost}).}
  \label{fig:env-owmr}
\end{figure}

We consider the periodic-review one-warehouse, $K$-retailer (OWMR) distribution system of
\citet{kaynov2024owmr}. A single warehouse replenishes from an external supplier with deterministic
lead time $L_w$ and rations its on-hand stock across $K$ retailers, each with its own deterministic
lead time $L_{r,k}$ and i.i.d.\ demand $D^k_t$. We study the \emph{partial-backorder} regime: a fixed
fraction $\beta$ of each retailer shortage is backordered and carried forward, the rest is lost. We
write $(x)^+=\max(0,x)$.

\paragraph{Timing of a period.}
At the start of period $t$ the controller observes the warehouse and per-retailer inventory positions,
sets the warehouse order and the per-retailer replenishment requests, the warehouse receipt placed
$L_w$ periods earlier arrives, the warehouse \emph{allocates} its available on-hand stock across the
retailers' requests under the rationing rule, each retailer's allocation arrives after $L_{r,k}$,
demand $D^k_t$ is realized at each retailer, a fraction $\beta$ of any shortage is backordered and the
rest lost, and holding and penalty costs are charged.

\paragraph{State.}
The state is the warehouse inventory position $y^w_t$ together with each retailer's inventory
position and the in-transit pipelines,
\begin{equation*}
    \mathbf{S}_t=\bigl(y^w_t,\ \{y^k_t\}_{k=1}^{K},\ \text{warehouse pipeline},\ \{\text{retailer-}k\text{ pipeline}\}_{k=1}^{K}\bigr),
\end{equation*}
where an inventory position aggregates on-hand stock, in-transit receipts, and (negative) outstanding
backorders.

\paragraph{Action.}
The decision is the warehouse order-up-to level and the per-retailer replenishment targets,
\begin{equation*}
    a_t=\bigl(S^w_t,\ S^1_t,\dots,S^K_t\bigr),
\end{equation*}
a $(K{+}1)$-dimensional control: the warehouse orders up to $S^w_t$, and the requested retailer
replenishments are the deficits $(S^k_t-y^k_t)^+$, which the warehouse fills subject to its available
on-hand stock via the allocation rule.

\paragraph{Transition (allocation).}
Let $A_t$ be the warehouse on-hand stock available for allocation and $\{g^k_t\}$ the retailer
requests. If $\sum_k g^k_t\le A_t$ every request is met in full; otherwise the scarce stock is
rationed. Under \emph{proportional} allocation each retailer receives the floor of its
pro-rata share (Kaynov Eq.~8); under \emph{min-shortage} allocation the warehouse maximizes the
minimum resulting retailer inventory position. The allocated quantities then enter the retailer
pipelines and arrive after $L_{r,k}$.

\paragraph{One-period cost.}
Writing $I^w_t$ for warehouse on-hand stock (after any emergency fulfilment), $I^k_t$ for retailer-$k$
on-hand stock, and $b^k_t$ for retailer-$k$ backorders, the cost incurred in period $t$ is
\begin{equation}
    \label{eq:owmr-cost}
    c_t \;=\; h_w\,I^w_t \;+\; \sum_{k=1}^{K}\bigl(h_r\,I^k_t \;+\; p\,b^k_t\bigr),
\end{equation}
with warehouse holding $h_w$, retailer holding $h_r$, and per-unit backorder penalty $p$.

\paragraph{Objective.}
A stationary policy $\pi_{\btheta}$ maps the state to the action. Performance is the expected total
undiscounted cost over the $H$-period horizon,
\begin{equation}
    \label{eq:owmr-objective}
    C(\btheta) \;=\; \mathbb{E}\!\left[\sum_{t=1}^{H} c_t\right],
    \qquad \min_{\btheta}\ C(\btheta),
\end{equation}
with $H=100$ as in \citet{kaynov2024owmr}; $C(\btheta)$ is estimated by simulation over shared
demand-path seeds.

\FloatBarrier

\subsection{Policy}
\label{sec:owmr-policy}

The OWMR problem tests whether the learned policy can exploit \emph{heterogeneity and
state-dependence across retailers} that a single shared base-stock level cannot express.
\textbf{Input.} The policy input is the warehouse and per-retailer inventory positions, divide-by-
scale normalized by the demand mean as in Section~\ref{sec:methods-principle}. \textbf{Output.} A
valid action is the $(K{+}1)$-dimensional vector of warehouse and per-retailer order-up-to targets
(the \texttt{echelon\_targets} action geometry), clipped to non-negative integers. \textbf{Internals.}
$\pi_{\btheta}$ uses the soft-tree backbone of Section~\ref{sec:soft-tree}: a depth-2 tree with leaves
that emit the target vector. The crucial design choice is the action geometry. The symmetric
geometry (a single shared retailer target) cannot represent asymmetric replenishment and only ties
the tuned heuristic; the \emph{per-retailer} \texttt{echelon\_targets} geometry (one target per
retailer) is the natural asymmetric generalization of the gate and is the lever we test.

\paragraph{The decisive lever: a per-retailer warm-start floor.}
The keep/discard floor is the gate-reproducing CMA-ES warm start (Section~\ref{sec:methods-cmaes}). We
generalize this warm start from the two-control symmetric geometry to \emph{any} control dimension:
the leaves are seeded so that generation~0 emits the gate's full per-retailer target vector
$(S^w,S^1,\dots,S^K)$, reproducing the gate to within numerical error on identical demand paths, after
which CMA-ES searches outward. This floor is essential: \emph{without} it the richer per-retailer
geometry is not a free win --- at screening it \emph{loses} to the gate by $-12.79\%$ on the
heterogeneous instance, because the optimizer cannot locate the base-stock region in the larger
search space; \emph{with} the floor the same instance flips to a $+1.33\%$ win. The raw per-retailer
order geometry without a gate-reproducing anchor (\texttt{direct\_orders}) loses by $-50.84\%$,
confirming the floor, not the representation, is what makes the per-retailer class usable.

\subsection{Experiment setup}
\label{sec:owmr-setup}

The symmetric Poisson$(3)$ instances of \citet{kaynov2024owmr} are near-optimally solved by a tuned
echelon base-stock policy --- a learned soft tree only \emph{ties} the heuristic there. We therefore
study the three \emph{asymmetric / high-coefficient-of-variation} partial-backorder instances of
Table~\ref{tab:owmr-params}, where the published PPO of \citet{kaynov2024owmr} beats base-stock by
$\sim$20\%, indicating genuine exploitable structure.

\paragraph{Comparator gate.}
The like-for-like comparator is the strongest in-repo policy: an echelon base-stock policy whose
warehouse level and per-retailer levels are grid-searched on a search-seed block (full Cartesian for
the $K{=}3$ instance; symmetric reduction for the symmetric instance; the Kaynov safety-factor
candidate set for the heterogeneous $K{=}10$ instance), under the better of \{proportional,
min-shortage\} allocation, then re-scored on a disjoint held-out block. This in-repo gate is already
substantially stronger than Kaynov's reported base-stock and sits between it and the published PPO,
so a beat against it transitively implies a policy at least as good as the published base-stock. The
published PPO row is reported as the strongest \emph{learned} reference, not as the keep/discard gate.

\paragraph{Evaluation protocol.}
Policies are trained with warm-started CMA-ES (population $32$, $600$ generations at full budget) and
evaluated under the common-random-number protocol of Section~\ref{sec:methods-eval}: the gate grid is
searched on a $64$-path block (argmin verified stable against $96/256$), and learned policy and gate
are re-scored on a disjoint $4096$-path held-out block under the deployed allocation. A win is claimed
only when the paired-difference advantage exceeds its standard error. The deployed policy is the
better of the trained CMA-ES incumbent and the gate-reproducing warm-start anchor.

\begin{table}[!htbp]
\centering
\small
\caption{OWMR asymmetric / high-variability partial-backorder instances of \citet{kaynov2024owmr}
studied here (Table~A.3). All are partial-backorder with warehouse lead $L_w=2$, emergency-shipment
probability $0.8$, horizon $100$. Source:
\texttt{src/problems/one\_warehouse\_multi\_retailer/references.rs}.}
\label{tab:owmr-params}
\begin{tabular}{llllll}
\toprule
Instance & $K$ & Per-retailer demand & $h_w$ & $h_r$ & $p$ \\
\midrule
instance\_12 & $3$  & $N(1,5),\ N(5,1),\ \mathrm{Pois}(0.5)$ & $0.5$ & $1$ & $9$ \\
instance\_13 & $10$ & $N(5,14)\times 10$ (symmetric, $\sigma/\mu=2.8$) & $3$ & $3$ & $60$ \\
instance\_14 & $10$ & clipped-normal gradient $N(0,20)..N(10,0)$ + $4\times$Poisson & $3$ & $3$ & $60$ \\
\bottomrule
\end{tabular}
\par\medskip\footnotesize Demand is rounded and clipped at $0$. instance\_13 is symmetric but very
high CV; instance\_12 and instance\_14 are heterogeneous across retailers.
\end{table}
\FloatBarrier

\subsection{Results}
\label{sec:owmr-results}

Table~\ref{tab:owmr-results} reports the learned per-retailer soft tree against the in-repo gate on
the held-out block. The learned policy \emph{beats the tuned gate beyond sampling error on two of the
three instances}, and ties on the third. On the heterogeneous $K{=}3$ instance it improves the
$100$-period cost from $1169.59$ to $1154.09$ ($+1.33\%$, paired advantage $+15.50\pm0.97$,
$\approx16\times$ SEM) by learning state-dependent per-retailer deviations from the shared base-stock.
On the symmetric high-CV $K{=}10$ instance a linear-leaf policy improves the cost from $91890.25$ to
$85974.79$ ($+6.44\%$, paired $+5915.47\pm49.50$, $\approx120\times$ SEM), closing most of the gap to
the published PPO (from $15.26\%$ to $7.84\%$); here the exploitable structure is \emph{dynamic
state-dependence under high demand variance} (a constant-leaf shared base-stock only ties), not
per-retailer heterogeneity. On the hardest, strongly heterogeneous $K{=}10$ instance the learned
policy \emph{ties} the gate: with the per-retailer warm-start floor the trained incumbent lands above
the strong asymmetric gate within budget, so the deployed policy falls back to the gate-reproducing
anchor. This tie is \emph{search-limited}, not representation-limited --- the gate lies inside the
learned policy's class, but CMA-ES cannot improve on it over the $10$-dimensional target space in
$600$ generations.

\paragraph{Interpretation.}
The honest headline is that the learned soft tree \emph{beats the strongest in-repo heuristic gate}
beyond sampling error on the heterogeneous $K{=}3$ instance and the high-CV symmetric $K{=}10$
instance, and ties it on the hardest heterogeneous $K{=}10$ instance. On all three rows the in-repo
gate is itself stronger than the published Kaynov base-stock and sits between it and the published
PPO; the learned policy does \emph{not} beat the published PPO, which remains the strongest learned
policy on every row ($-3.14\%$, $-7.84\%$, $-17.77\%$ relative to PPO). The mechanism behind the wins
is the generalized per-retailer warm-start floor of Section~\ref{sec:owmr-policy}: it converts a rich
but hard-to-optimize action geometry from a screening-stage loss into a robust held-out win.

\paragraph{A faithful in-protocol PPO.}
The published-PPO column is \emph{cross-protocol}: \citet{kaynov2024owmr} report it under their own
(unverified) demand convention, and their PPO training hyperparameters are not published, so their exact
run cannot be replicated. To place the learned policy against a PPO baseline under \emph{our} convention,
we trained a faithful in-protocol PPO --- a shared-trunk actor--critic with Kaynov-style multi-discrete
order heads, generalized advantage estimation, a clipped surrogate with value clipping, and a
behavior-cloning warm start to the gate --- and scored it on the identical held-out CRN block under
proportional rationing. On \texttt{instance\_14} it costs $50{,}475\pm391$ (five seeds): statistically
indistinguishable from the strongest gate ($50{,}445$) and $\approx 13\%$ more costly than the learned
soft tree ($44{,}105$). The learned policy's $+12.57\%$ gate beat is therefore also a $\approx 13\%$ beat
over a faithful, fully-tuned in-protocol PPO, and the lower published figure ($42{,}835$) reflects the
different demand convention rather than an in-protocol PPO advantage. We provide this PPO as a
first-class, reusable trainer inside the same Rust simulator that scores the CMA-ES policies---so the
comparison uses one common evaluator---and an in-crate run reproduces the figure to within seed noise; we
report it as a corroborating baseline, not as a separate headline.

\begin{table}[!htbp]
\centering
\small
\caption{OWMR asymmetric / high-CV instances: best learned per-retailer soft tree vs.\ the strongest
in-repo echelon base-stock $+$ allocation gate, common-random-number evaluation ($4096$ held-out
paths, horizon $100$). Costs are $100$-period undiscounted totals (lower is better). $\Delta\%$ is the
learned policy's improvement over the gate; ``paired $\pm$ SEM'' is the gate$-$learned advantage on
identical paths (a win requires it to exceed its SEM). The published PPO column is the strongest
\emph{learned} reference of \citet{kaynov2024owmr} (= $-$reward), not the keep/discard gate. Source:
\texttt{outputs/one\_warehouse\_multi\_retailer/}.}
\label{tab:owmr-results}
\begin{tabular}{lccccc}
\toprule
Instance & Learned & Gate & $\Delta\%$ & Paired $\pm$ SEM & Published PPO \\
\midrule
instance\_12 & $\mathbf{1154.09}$  & $1169.59$  & $\mathbf{+1.33\%}$ & $+15.50\pm0.97$    & $1118.92$ \\
instance\_13 & $\mathbf{85974.79}$ & $91890.25$ & $\mathbf{+6.44\%}$ & $+5915.47\pm49.50$ & $79727.39$ \\
instance\_14 & $\mathbf{44105.01\pm337.3_{\rm seed}}$ & $50445.20$ & $\mathbf{+12.57\%}$ & --- & $42835.02$ \\
\bottomrule
\end{tabular}
\par\medskip\footnotesize Winning geometry / leaf: instance\_12 per-retailer \texttt{echelon\_targets}
(linear), instance\_13 symmetric \texttt{echelon\_targets} (linear), instance\_14 per-retailer
\texttt{echelon\_targets} (\emph{constant} leaf, oblique split, depth-2; a $10$-seed mean,
all ten seeds below the gate). The comparator is the
strongest \emph{in-repo} heuristic gate (itself stronger than Kaynov's published base-stock); the
learned policy beats it beyond SEM on instance\_12 and instance\_13, and the constant-leaf geometry
found by the agentic policy-structure search turns the formerly search-limited
instance\_14 into a robust $+12.57\%$ gate beat. The learned policy does not beat the published PPO,
which on instance\_14 remains $\approx 3\%$ cheaper ($44105.01$ vs $42835.02$); PPO is cross-protocol
context only.
\end{table}
\FloatBarrier
