\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[numbers]{natbib}

\title{Fixed-Order-Cost Lost-Sales Control with CMA-ES-Trained Policy Families}
\author{Working manuscript}
\date{\today}

\begin{document}
\maketitle

\section{Problem description}

We study a single-item, periodic-review lost-sales inventory problem with positive lead time and a
fixed setup cost for every positive order. In each period $t$, the controller observes the current
inventory state and chooses an integer order quantity $q_t \in \{0,1,\ldots,\bar q\}$. Demand is
stochastic, unmet demand is lost, and the one-period cost contains three components:
\begin{enumerate}
    \item inventory holding cost,
    \item lost-sales penalty,
    \item fixed ordering cost $K \mathbf{1}\{q_t > 0\}$.
\end{enumerate}

Relative to the classical lost-sales problem, the fixed ordering cost changes the economics of the
decision problem substantially. The controller must not only determine how much to order, but also
whether ordering at all is worth paying the setup cost. This creates a natural no-order region and
a separate positive-order region, which in turn makes the action parameterization of learned
policies much more important than in the fixed-cost-free benchmark.

\section{Reference instance and benchmark grid}

Our immediate focus is the canonical fixed-cost instance
\texttt{lit\_pois\_mu5\_l4\_p4\_k5}, which has:
\begin{itemize}
    \item lead time $L=4$,
    \item shortage cost $p=4$,
    \item holding cost $h=1$,
    \item fixed ordering cost $K=5$,
    \item Poisson demand with mean $\mu=5$.
\end{itemize}

This instance is part of the benchmark family of \citet{Bijvank2015ParametricPolicies}. In the
codebase, the first literature-aligned subset grid is
\texttt{literature\_subset\_poisson\_mu5}, which spans:
\[
L \in \{1,2,3,4\}, \qquad p \in \{4,19\}, \qquad K \in \{5,25\},
\]
with Poisson demand of mean $5$ and holding cost $1$ throughout.

The literature provides the benchmark family and the heuristic classes, but it does not publish a
clean exact per-instance table for our canonical instance. Therefore, the numerical values reported
here are repo-native benchmark values computed on that published instance family.

\section{Heuristic baselines}

Following \citet{Bijvank2015ParametricPolicies}, we benchmark against three classical parametric
policy classes:
\begin{itemize}
    \item $(s,S)$,
    \item $(s,nQ)$,
    \item modified $(s,S,q)$.
\end{itemize}

For the canonical instance, heuristic parameters are obtained by explicit search over bounded
parameter domains using the same simulator as for the learned policies. The current best heuristic
on the canonical instance is modified $(s,S,q)$.

\section{Implementation validation}

Before using the fixed-order-cost simulator and heuristic benchmark pipeline for new learned-policy
experiments, we validate the implementation against a published example from
\citet{Bijvank2015ParametricPolicies}. Their Table~1 reports a small Poisson-demand instance with
lead time $L=2$, shortage cost $p=14$, fixed ordering cost $K=5$, holding cost $h=1$, and demand
mean $\mu=5$. For that instance, the paper reports the exact optimal cost and the best parameters
for the three heuristic policy classes that we implement.

\begin{table}[t]
\centering
\begin{tabular}{llll}
\toprule
Policy & Reported parameters & Published cost & Our reproduced cost \\
\midrule
optimal & --- & 11.46 & --- \\
$(s,S)$ & $(17,23)$ & 11.62 & 11.6217 \\
$(s,nQ)$ & $(17,7)$ & 11.56 & 11.5589 \\
modified $(s,S,q)$ & $(17,23,7)$ & 11.50 & 11.5012 \\
\bottomrule
\end{tabular}
\caption{Published fixed-order-cost lost-sales validation example from \citet{Bijvank2015ParametricPolicies} and our reproduced long-run costs for the reported heuristic parameters.}
\label{tab:fixed-cost-validation}
\end{table}

Table~\ref{tab:fixed-cost-validation} shows that our long-run simulator reproduces the published
heuristic costs to within simulation precision. In particular, the $(s,S)$ policy matches both the
reported parameterization and the reported cost essentially exactly, while the $(s,nQ)$ and
modified $(s,S,q)$ costs differ from the published values only in the third decimal place. This is
strong evidence that the fixed-order-cost environment dynamics, cost accounting, and heuristic
policy implementations are correct.

One subtlety is that simulation-based heuristic search over a finite sampled demand path need not
return a unique optimizer when several parameterizations are nearly tied. On this validation
instance, direct long-run evaluation of the published parameters matches the paper very closely,
which is the relevant correctness check. Therefore, in the experiments below we treat the
fixed-order-cost heuristic stack as externally validated against the literature, and we report both
the selected heuristic parameters and their long-run reevaluated costs.

We also ran the full learned-policy suite on this same known-optimum instance. Because the exact
optimal cost is available here, this is the cleanest place to compare learned policies to the true
benchmark rather than only to the best heuristic.

\begin{table}[t]
\centering
\caption{Known-optimum validation benchmark on the published Table~1 instance
\texttt{bijvank2015\_table1\_l2\_p14\_k5}. Since the exact optimum is available, the main
comparison is gap to that optimum.}
\label{tab:fixed-cost-validation-suite}
\begin{tabular}{lrr}
\toprule
Policy family & Mean cost & Gap vs.\ optimum \\
\midrule
\multicolumn{3}{l}{\textit{Exact and heuristic baselines}} \\
optimal & 11.4600 & 0.0000 \\
$(s,S)$ & 11.6192 & 0.1592 \\
$(s,nQ)$ & 11.6317 & 0.1717 \\
modified $(s,S,q)$ & 11.5038 & 0.0438 \\
\midrule
\multicolumn{3}{l}{\textit{Learned policy families}} \\
Linear, categorical quantity & 12.5219 & 1.0619 \\
Linear, gated ordinal quantity & 11.5260 & 0.0660 \\
NN, categorical quantity & 12.2939 & 0.8339 \\
NN, gated ordinal quantity & 11.5436 & 0.0836 \\
Soft tree, oblique depth-2 linear leaf & 12.1070 & 0.6470 \\
Soft tree, oblique depth-1 linear leaf & 12.1252 & 0.6652 \\
\bottomrule
\end{tabular}
\end{table}

Table~\ref{tab:fixed-cost-validation-suite} shows that the best learned policy on the published
instance is the linear policy with the gated ordinal quantity head, at cost $11.5260$, only
$0.0660$ above the exact optimum and $0.0222$ above the best heuristic. The neural gated ordinal
policy is close behind at $11.5436$. Thus the known-optimum validation instance supports the same
qualitative conclusion as the canonical benchmark: the gated ordinal head is the only learned
parameterization that gets close to the correct order/no-order structure of the fixed-cost problem.

\section{Learned policy families}

We compare three policy backbones:
\begin{itemize}
    \item a linear policy,
    \item a neural network policy,
    \item a soft oblique decision tree with linear leaves.
\end{itemize}

The key modeling axis is the action head:
\begin{itemize}
    \item \textbf{categorical quantity}: directly chooses one order quantity by categorical
    argmax;
    \item \textbf{gated ordinal quantity}: separates the tendency to order from the quantity score,
    which is much better aligned with the economics of a fixed setup cost.
\end{itemize}

The central hypothesis for this problem is that the action parameterization matters more than the
backbone. If that hypothesis is correct, then moving from categorical quantity to gated ordinal
quantity should improve both the linear and neural policies materially.

The categorical quantity baseline is the same flat action parameterization that worked well for the
plain lost-sales problem: the policy emits $\bar q+1$ logits and chooses a single order quantity by
argmax:
\[
a_t = \arg\max_{a \in \{0,1,\ldots,\bar q\}} z_a(s_t),
\]
where $z_0(s_t),\ldots,z_{\bar q}(s_t)$ are the logits emitted by the policy. The fixed-cost
results show that this flat action space is a poor fit once a setup cost creates an explicit
no-order region.

For the gated ordinal quantity head, the policy backbone produces one scalar gate score $g(s_t)$
and $\bar q$ ordinal scores $o_1(s_t),\ldots,o_{\bar q}(s_t)$. We then define
\[
\hat m(s_t) = \sum_{k=1}^{\bar q} \sigma(o_k(s_t)),
\qquad
a_t = \mathrm{round}\!\left(\sigma(g(s_t)) \, \hat m(s_t)\right),
\]
where $\sigma(\cdot)$ is the logistic sigmoid. The gate $\sigma(g(s_t))$ represents the tendency
to place any positive order, while $\hat m(s_t)$ represents the magnitude of the positive-order
decision. This is still a general learned policy, not a hand-coded $(s,S)$-type rule, but it
searches in a control space that is much closer to the economics of the problem.

The soft-tree family uses a different backbone but the same underlying idea of a structured action
map. Each internal node computes a soft split
\[
p_j(s_t) = \sigma(w_j^\top s_t - b_j),
\]
which induces path probabilities $\pi_\ell(s_t)$ over the leaves. Each leaf $\ell$ emits a small
linear quantity model
\[
q_\ell(s_t) = \alpha_\ell^\top s_t + \beta_\ell,
\]
and the final action is the rounded weighted mixture
\[
a_t = \mathrm{round}\!\left(\sum_{\ell} \pi_\ell(s_t)\, q_\ell(s_t)\right).
\]
The tree does not commit to a hard branch during training; instead, soft splits induce path
probabilities over leaf models, which keeps the policy differentiable-free yet smooth enough for
stable CMA-ES search. The depth-1 and depth-2 oblique trees with linear leaves form the strongest
tree baselines in the current benchmark.

\section{Benchmark protocol}

All learned policies in the current canonical benchmark use the same CMA-ES training budget:
\begin{itemize}
    \item $5000$ training episodes,
    \item population size $50$,
    \item training horizon $2000$,
    \item evaluation horizon $10^6$,
    \item $10$ evaluation seeds,
    \item warm-up ratio $0.2$.
\end{itemize}

The canonical benchmark runner is
\texttt{scripts/lost\_sales\_fixed\_order\_cost/benchmark\_canonical\_suite.py}. The canonical
result table is generated directly from the benchmark JSON, so the manuscript table remains
synchronized with the code and experiment outputs.

\section{Canonical results}

\begin{table}[t]
\centering
\caption{Canonical fixed-order-cost lost-sales benchmark for $L=4$, $p=4$, $K=5$, and Poisson demand with mean $5$. All learned policies use $5000$ CMA-ES iterations, population size $50$, training horizon $2000$, and long-run evaluation over $10$ seeds with horizon $10^6$.}
\label{tab:fixed-cost-canonical}
\begin{tabular}{lrr}
\toprule
Policy family & Mean cost & Gap vs.\ best heuristic \\
\midrule
\multicolumn{3}{l}{\textit{Heuristic baselines}} \\
$(s,S)$ & 9.3715 & 0.1971 \\
$(s,nQ)$ & 9.1810 & 0.0066 \\
modified $(s,S,q)$ & 9.1744 & 0.0000 \\
\midrule
\multicolumn{3}{l}{\textit{Learned policy families}} \\
Linear, categorical quantity & 10.2730 & 1.0986 \\
Linear, gated ordinal quantity & 8.7688 & -0.4056 \\
NN, categorical quantity$^{\dagger}$ & 10.2730 & 1.0986 \\
NN, gated ordinal quantity & 8.7328 & -0.4415 \\
Soft tree, oblique depth-2 linear leaf & 8.7742 & -0.4002 \\
Soft tree, oblique depth-1 linear leaf & 8.7785 & -0.3959 \\
\bottomrule
\end{tabular}
\par\medskip\footnotesize{$^{\dagger}$ The current NN categorical-quantity baseline matches the linear categorical baseline exactly and should be re-verified before publication claims rely on it.}
\end{table}

The best heuristic baseline is modified $(s,S,q)$ at cost $9.1744$. The strongest trusted learned
policy is the neural network with the gated ordinal quantity head, at cost $8.7328$. The linear
policy with the same head achieves $8.7688$, which places it essentially on top of the two best
tree variants at $8.7742$ and $8.7785$.

These results support a clear conclusion: on the fixed-order-cost lost-sales problem, the action
head is the dominant architectural choice. The plain categorical quantity head performs poorly,
whereas the gated ordinal quantity head makes even the linear policy competitive with the more
expressive tree and neural policies.

One baseline should still be treated cautiously: the current
\texttt{nn\_categorical\_quantity} run matches the linear categorical baseline exactly. We retain
that row in the benchmark table for completeness, but it should be re-verified before publication
claims rely on it.

\section{Full-grid results}

To move beyond the canonical instance, we also aggregate performance over the full
\texttt{literature\_subset\_poisson\_mu5} grid. For each of the $16$ instances in that grid, we
search the three heuristic families, train the six learned-policy families under the common CMA-ES
budget, and compare each learned policy against the best heuristic found on that same instance.
The aggregate metric below is the mean relative gap to that per-instance best heuristic, together
with the number of instances on which the learned policy beats it.

\begin{table}[t]
\centering
\caption{Aggregate fixed-order-cost lost-sales results over the literature-aligned subset grid with $L \in \{1,2,3,4\}$, $p \in \{4,19\}$, $K \in \{5,25\}$, and Poisson demand with mean $5$. Each row reports the mean relative gap to the best heuristic found on each instance and the number of instances on which the learned policy beats that best heuristic. Lower is better.}
\label{tab:fixed-cost-full-grid}
\begin{tabular}{lrrl}
\toprule
Policy family & Mean relative gap (\%) & Better than best heuristic & Status \\
\midrule
Linear, categorical quantity & 692842.158 & 0/16 & trusted \\
Linear, gated ordinal quantity & 5.984 & 8/16 & trusted \\
NN, categorical quantity$^{\dagger}$ & 15.475 & 2/16 & provisional \\
NN, gated ordinal quantity & -0.809 & 11/16 & trusted \\
Soft tree, oblique depth-2 linear leaf & 16.833 & 0/16 & trusted \\
Soft tree, oblique depth-1 linear leaf & 13.960 & 1/16 & trusted \\
\bottomrule
\end{tabular}
\par\medskip\footnotesize{Best trusted aggregate row: NN, gated ordinal quantity with mean relative gap -0.809\% and wins on 11/16 instances. $^{\dagger}$ The current NN categorical-quantity baseline remains provisional and should be re-verified before publication claims rely on it.}
\end{table}

Like the canonical table, this full-grid artifact is generated directly from the benchmark JSON so
the manuscript stays synchronized with the experiment outputs.

To make the instance-level picture explicit in the same style as the plain lost-sales paper,
Tables~\ref{tab:fixed-cost-k5-slice} and~\ref{tab:fixed-cost-k25-slice} report the actual mean
costs on the $L \in \{2,4\}$ slice of the fixed-cost grid. These are the costs from the full-grid
run itself, so the canonical $L=4$, $p=4$, $K=5$ entry need not match the separate canonical-suite
table exactly. The Geometric columns are left blank for now because we have not yet run the
aligned fixed-cost Geometric benchmark.

\begin{table}[t]
\caption{Average cost of policy function approximators on the fixed-order-cost lost-sales problem
with fixed order cost $K=5$ and lead times in $\{2,4\}$. Geometric columns are left empty because
those runs have not yet been executed for the fixed-cost benchmark.}
\label{tab:fixed-cost-k5-slice}
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|c|c||c|c|}
\hline
 & & \multicolumn{2}{c||}{Lead Time = 2} & \multicolumn{2}{c|}{Lead Time = 4} \\
Shortage & Policy & Geometric & Poisson & Geometric & Poisson \\
Cost & & & & & \\ \hline
4 & $(s,S)$ &  & 8.8695 &  & 9.5619 \\
  & $(s,nQ)$ &  & 8.8071 &  & 9.2123 \\
  & modified $(s,S,q)$ &  & 8.8021 &  & 9.2032 \\
  & Linear, categorical &  & 9.6582 &  & 10.2730 \\
  & Linear, gated ordinal &  & 8.5974 &  & 8.7723 \\
  & NN, categorical &  & 9.4855 &  & 10.2730 \\
  & NN, gated ordinal &  & 8.5173 &  & 8.7929 \\
  & Soft tree, depth-2 &  & 9.2744 &  & 9.7519 \\
  & Soft tree, depth-1 &  & 9.4218 &  & 9.7530 \\ \hline
19 & $(s,S)$ &  & 12.2996 &  & 13.7219 \\
   & $(s,nQ)$ &  & 12.2365 &  & 13.6427 \\
   & modified $(s,S,q)$ &  & 12.1976 &  & 13.5483 \\
   & Linear, categorical &  & 13.2537 &  & 14.9028 \\
   & Linear, gated ordinal &  & 12.1783 &  & 13.6092 \\
   & NN, categorical &  & 12.9253 &  & 14.3474 \\
   & NN, gated ordinal &  & 12.2043 &  & 13.4420 \\
   & Soft tree, depth-2 &  & 12.7438 &  & 13.9892 \\
   & Soft tree, depth-1 &  & 12.7433 &  & 14.0448 \\ \hline
\end{tabular}%
}
\end{table}

\begin{table}[t]
\caption{Average cost of policy function approximators on the fixed-order-cost lost-sales problem
with fixed order cost $K=25$ and lead times in $\{2,4\}$. Geometric columns are left empty because
those runs have not yet been executed for the fixed-cost benchmark.}
\label{tab:fixed-cost-k25-slice}
\centering
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|l|c|c||c|c|}
\hline
 & & \multicolumn{2}{c||}{Lead Time = 2} & \multicolumn{2}{c|}{Lead Time = 4} \\
Shortage & Policy & Geometric & Poisson & Geometric & Poisson \\
Cost & & & & & \\ \hline
4 & $(s,S)$ &  & 15.4044 &  & 15.9471 \\
  & $(s,nQ)$ &  & 15.3194 &  & 15.8247 \\
  & modified $(s,S,q)$ &  & 15.3319 &  & 15.8625 \\
  & Linear, categorical &  & 15.3194 &  & 19.9952 \\
  & Linear, gated ordinal &  & 19.9952 &  & 15.6355 \\
  & NN, categorical &  & 15.3410 &  & 15.5593 \\
  & NN, gated ordinal &  & 15.3204 &  & 15.5705 \\
  & Soft tree, depth-2 &  & 19.9952 &  & 19.9952 \\
  & Soft tree, depth-1 &  & 19.9952 &  & 19.9952 \\ \hline
19 & $(s,S)$ &  & 19.5629 &  & 21.1835 \\
   & $(s,nQ)$ &  & 19.6401 &  & 21.0192 \\
   & modified $(s,S,q)$ &  & 19.6538 &  & 20.9927 \\
   & Linear, categorical &  & 34.3710 &  & 34.4796 \\
   & Linear, gated ordinal &  & 19.5581 &  & 21.7506 \\
   & NN, categorical &  & 19.6040 &  & 21.1661 \\
   & NN, gated ordinal &  & 19.5766 &  & 20.7843 \\
   & Soft tree, depth-2 &  & 25.0137 &  & 26.4984 \\
   & Soft tree, depth-1 &  & 19.5571 &  & 26.4972 \\ \hline
\end{tabular}%
}
\end{table}

The completed $16$-instance run strengthens the canonical-instance conclusion. Among the trusted
families, the neural network with the gated ordinal quantity head is best in aggregate, with mean
relative gap $-0.809\%$ versus the best heuristic found on each instance and wins on $11/16$
instances. The linear policy with the same gated ordinal head is the next strongest trusted family
at $5.984\%$ mean relative gap and wins on $8/16$ instances. Thus the main structural result
survives the move from one canonical instance to the full literature-aligned grid: separating the
decision to order from the positive-order quantity score matters more than moving from a linear
backbone to a more expressive one.

The instance-level pattern is also informative. On the lower setup-cost half of the grid
($K=5$), every instance is won by one of the gated ordinal policies, split $5$ wins for the neural
policy and $3$ wins for the linear policy. At lead time $L=3$, the neural gated ordinal policy is
the best learned family on all four grid instances. By contrast, the provisional
\texttt{nn\_categorical\_quantity} row only looks competitive on two $p=4$, $K=25$ cases, so we do
not use it for publication-critical conclusions.

Two caveats from the completed run should remain visible. First, the flat linear categorical
quantity policy is not merely weak on average; it fails catastrophically on some instances. The
worst case is \texttt{lit\_pois\_mu5\_l3\_p19\_k5}, where its long-run mean cost is
$1{,}440{,}013.70$ against a best-heuristic benchmark of $12.9904$, which drives its aggregate
mean relative gap up to $692842.2\%$. Second, on the three $p=4$, $K=25$ instances with
$L \in \{1,2,3\}$, the linear gated policy and both soft-tree baselines all collapse to the same
long-run cost $19.995229$, producing relative gaps between roughly $29\%$ and $33\%$. That looks
more like a repeated failure mode than ordinary simulation noise, so those rows should be treated
as a debugging target rather than a fine-grained modeling conclusion.

With those caveats stated explicitly, the robust empirical takeaway is still clear: the gated
ordinal action head is the only design change that consistently transfers across the fixed-cost
grid, and the neural gated ordinal policy is currently the strongest trusted learned policy family
for this problem class.

\section{Current paper scope}

At this point the fixed-order-cost lost-sales problem is the first extension problem that is mature
enough for a stand-alone empirical section:
\begin{itemize}
    \item the model is implemented and documented,
    \item the benchmark family and canonical instance are explicit,
    \item the classical heuristics are in place,
    \item the learned policy matrix has a stable paper-like benchmark protocol,
    \item the manuscript can consume both canonical and full-grid benchmark artifacts directly.
\end{itemize}

The next step is to consolidate the fixed-cost section with the dual-sourcing and multi-echelon
problem sections in the broader manuscript.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
