\documentclass[11pt]{article}

\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{booktabs}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[numbers]{natbib}

\title{Inventory Control Policy Benchmarks for Selected Inventory Problems}
\author{Working manuscript}
\date{\today}

\begin{document}
\maketitle

\section{Problem focus}

This note records the current architecture study for the $L=4$, $p=4$ lost-sales family. The
central question is not whether one policy class is universally optimal, but which policy
parameterizations expose a replenishment rule that CMA-ES can search reliably.

The main numerical study below is still the lost-sales benchmark suite. We additionally record
three single-instance Rust-side benchmarks for perishable inventory, random-yield inventory, and
joint replenishment, using the same trained soft-tree policy family and the current
literature-verification rule: when public benchmark numbers are available, we report against them;
otherwise we report against our own exact solver and explicitly mark the instance as not verified
against a published table.

We study two closely related problems:
\begin{itemize}
    \item \textbf{vanilla lost sales}: no fixed setup cost,
    \item \textbf{fixed-cost lost sales}: a positive fixed ordering cost $K$ is charged whenever
    $q_t > 0$.
\end{itemize}

In both cases the controller observes the current inventory state and chooses an integer order
quantity $q_t \in \mathbb{Z}_{\ge 0}$. Demand is stochastic and unmet demand is lost. The
fixed-cost version adds the term $K\mathbf{1}\{q_t>0\}$ to the one-period cost, which makes the
order/no-order boundary much more important. Some policy families still introduce a finite
decoder-side cap $\bar q$, but that cap is now a policy design choice rather than an environment
restriction.

\section{Current benchmark instances and protocol}

The current single-instance study uses:
\begin{itemize}
    \item vanilla instance \texttt{vanilla\_l4\_p4\_poisson5},
    \item fixed-cost instance \texttt{lit\_pois\_mu5\_l4\_p4\_k5}.
\end{itemize}

Their shared core parameters are:
\[
L = 4, \qquad p = 4, \qquad h = 1, \qquad \mu = 5.
\]
The fixed-cost instance additionally uses $K=5$.

All current runs in this note use the same training budget:
\begin{itemize}
    \item $2000$ CMA-ES iterations,
    \item population size $64$,
    \item training horizon $2000$,
    \item evaluation horizon $10^6$,
    \item $10$ evaluation seeds,
    \item base random seed $42$ for training and heuristic search; reported mean and standard
    deviation values are evaluated over the consecutive seed set
    $\{42,43,\ldots,51\}$.
\end{itemize}

For the demand-robustness sweep we use four mean-preserving demand families:
\begin{itemize}
    \item Poisson demand with mean $5$,
    \item Geometric demand with mean $5$,
    \item positively correlated two-state Markov-modulated Poisson demand with
    $(\lambda_{\mathrm{low}}, \lambda_{\mathrm{high}})=(3,7)$ and $p_{00}=p_{11}=0.9$,
    \item negatively correlated two-state Markov-modulated Poisson demand with
    $(\lambda_{\mathrm{low}}, \lambda_{\mathrm{high}})=(3,7)$ and $p_{00}=p_{11}=0.1$.
\end{itemize}

For the two-state Markov-modulated Poisson cases, the latent regime $S_t \in \{\mathrm{low},
\mathrm{high}\}$ evolves as a two-state Markov chain with transition matrix
\[
P =
\begin{bmatrix}
p_{00} & 1-p_{00} \\
1-p_{11} & p_{11}
\end{bmatrix},
\]
and demand is then sampled as
\[
D_t \mid S_t=\mathrm{low} \sim \mathrm{Poisson}(\lambda_{\mathrm{low}}),
\qquad
D_t \mid S_t=\mathrm{high} \sim \mathrm{Poisson}(\lambda_{\mathrm{high}}).
\]
Because the reported cases use $p_{00}=p_{11}$, the stationary regime probabilities are
$\Pr(S_t=\mathrm{low})=\Pr(S_t=\mathrm{high})=1/2$, so both correlated variants preserve the same
unconditional mean demand:
\[
\mathbb{E}[D_t] = \tfrac{1}{2}\lambda_{\mathrm{low}} + \tfrac{1}{2}\lambda_{\mathrm{high}}
= \tfrac{1}{2}(3+7)=5.
\]
The sign and magnitude of temporal dependence come from the lag-one regime eigenvalue
\[
\rho_{\mathrm{regime}} = p_{00}+p_{11}-1.
\]
In the positive-correlation case this equals $0.8$, so the regime is persistent and the expected
run length in either state is $1/(1-0.9)=10$ periods. In the negative-correlation case it equals
$-0.8$, so the latent regime is strongly alternating. After Poisson observation noise is added,
the implemented lag-one demand autocorrelations become
\[
\mathrm{Corr}(D_t,D_{t+1})
= \frac{\pi_{\mathrm{low}}\pi_{\mathrm{high}}(\lambda_{\mathrm{high}}-\lambda_{\mathrm{low}})^2
(p_{00}+p_{11}-1)}{\mathrm{Var}(D_t)}
= \pm \frac{3.2}{9}
\approx \pm 0.356,
\]
with the plus sign for the persistent case and the minus sign for the alternating case. These are
the first-lag correlations checked explicitly in the Rust demand tests. More generally, the
implemented lag-$k$ demand autocovariance is proportional to $(p_{00}+p_{11}-1)^k$, so the
positive case produces geometrically decaying positive dependence and the negative case produces
geometrically decaying sign-alternating dependence. By contrast, the Poisson and Geometric i.i.d.
families have zero autocorrelation at every positive lag.

\section{Policy families and action equations}

The current study is intentionally narrow: the reported experiments use linear policies and soft
trees on the nonnegative scalar action space
\[
a_t \in \mathbb{Z}_{\ge 0}.
\]
Let $x_t \in \mathbb{R}^d$ denote the state feature vector. Some decoders additionally introduce a
finite policy-side cap $\bar q$ and therefore operate on the bounded control set
\[
\mathcal{A}_{\bar q} = \{0,1,\ldots,\bar q\},
\qquad
\mathcal{A}_{\bar q}^{+} = \{1,\ldots,\bar q\}.
\]
For uncapped nonnegative heads we write
\[
\mathcal{A}_{\infty} = \mathbb{Z}_{\ge 0}.
\]
The code implements a \emph{backbone} plus a \emph{decoder}:
\begin{itemize}
    \item a linear backbone emits logits by an affine map,
    \item a one-hidden-layer neural backbone emits logits through an MLP,
    \item a soft tree emits split probabilities and leaf-local quantity predictions.
\end{itemize}

For dense policies, let $u_\theta(x_t)$ denote the backbone output. In the linear case,
\[
u_\theta(x_t) = W x_t + b.
\]
In the one-hidden-layer neural case with width $H$ and activation $\phi$,
\[
u_\theta(x_t) = V \phi(Ux_t + c) + e.
\]
The decoder formulas below are shared by both dense backbones; only the way the latent outputs are
produced changes. For scalar controls we write $\Pi_{\mathcal{A}}(z)$ for nearest-integer
projection of a real-valued proposal $z$ onto the discrete action set $\mathcal{A}$.

We study the following action parameterizations:

\paragraph{Categorical quantity.}
The policy emits $\bar q+1$ logits and chooses
\[
a_t = \arg\max_{a \in \mathcal{A}_{\bar q}} u_a(x_t).
\]
This head is inherently cap-dependent.

\paragraph{Sigmoid direct quantity.}
The policy emits one scalar and maps it as
\[
a_t = \mathrm{round}\!\left(\bar q \, \sigma(u(x_t))\right).
\]
This is the code path named \texttt{sigmoid\_direct\_quantity}: a one-head bounded S-curve map.

\paragraph{Direct quantity.}
The policy emits one scalar and maps it as
\[
a_t = \Pi_{\mathcal{A}_{\infty}}\!\left(\mathrm{softplus}(u(x_t))\right).
\]
This is the code path named \texttt{direct\_quantity}: a one-head positive-part map with no
decoder-side upper bound.

\paragraph{Capped direct quantity.}
The policy emits one scalar and maps it as
\[
a_t = \Pi_{\mathcal{A}_{\bar q}}\!\left(\mathrm{softplus}(u(x_t))\right).
\]
This is the code path named \texttt{capped\_direct\_quantity}. It is the same softplus quantity
map as above, but with a decoder-side cap entering only through the final projection.

\paragraph{Soft-gated sigmoid direct quantity.}
The policy emits two latent outputs, a gate score $g(x_t)$ and a quantity score $q(x_t)$, then
sets
\[
a_t = \mathrm{round}\!\left(\sigma(g(x_t)) \, \sigma(q(x_t)) \, \bar q\right).
\]
This is the code path named \texttt{gated\_sigmoid\_direct\_quantity}.

\paragraph{Soft-gated direct quantity.}
The policy emits $g(x_t)$ and $q(x_t)$, then sets
\[
a_t = \Pi_{\mathcal{A}_{\bar q}}\!\left(\sigma(g(x_t)) \, \mathrm{softplus}(q(x_t))\right).
\]
This is the code path named \texttt{soft\_gated\_direct\_quantity}.

\paragraph{Hard-gated direct quantity.}
The policy emits $g(x_t)$ and $q(x_t)$, and uses
\[
a_t =
\begin{cases}
0, & \sigma(g(x_t)) < 0.5, \\
\Pi_{\mathcal{A}_{\bar q}^{+}}\!\left(\mathrm{softplus}(q(x_t))\right),
& \sigma(g(x_t)) \ge 0.5.
\end{cases}
\]
This is the code path named \texttt{hard\_gated\_direct\_quantity}.

\paragraph{Soft trees with linear leaves.}
A soft tree computes path probabilities $\pi_\ell(x_t)$ over leaves, and each leaf emits a local
linear quantity model
\[
q_\ell(x_t) = \mathrm{softplus}(\alpha_\ell^\top x_t + \beta_\ell).
\]
The final action is the rounded mixture
\[
a_t = \Pi_{\mathcal{A}_{\bar q}}\!\left(\sum_\ell \pi_\ell(x_t) q_\ell(x_t)\right).
\]
This is the active \texttt{soft\_tree\_depth1\_linear\_leaf} /
\texttt{soft\_tree\_depth2\_linear\_leaf} family. In code, the split network is always oblique in
the reported runs and the scalar leaf is exactly a softplus-transformed affine quantity rule.

The policy family that is \emph{not} part of the active benchmark surface is the raw unbounded
direct head
\[
a_t = \Pi_{\mathcal{A}_{\bar q}}\!\left(u(x_t)\right),
\]
which repeatedly converged to weak basins in earlier runs and is now treated as a negative ablation
rather than a serious candidate.

\subsection{Parameter counts}

The policy families discussed in this note fall into three groups: linear dense policies,
one-hidden-layer neural dense policies, and soft trees. For the active $L=4$ pipeline state we
have $d=4$ and therefore $d+1=5$. The bounded heads use $\bar q=20$ in the canonical vanilla
instance and $\bar q=50$ in the canonical fixed-cost instance. The current NN registry entries use
a single hidden layer of width $H=50$ with SELU activation. We exclude the ordinal decoders from
the active rerun surface below because their output dimension, and hence their parameter count,
still scales directly with $\bar q$.

\begin{table}[t]
\centering
\small
\caption{Trainable parameter counts for the linear scalar-quantity policy families. The last
column shows the concrete $L=4$ counts, reported as vanilla / fixed-cost when the decoder-side
cap differs across the two canonical instances.}
\label{tab:linear-parameter-counts}
\begin{tabular}{|l|c|c|}
\hline
Policy family & Trainable parameters & Count at $L=4$ \\
\hline
Linear categorical quantity & $(d+1)(\bar q+1)$ & $105 / 255$ \\
Linear sigmoid direct quantity & $d+1$ & $5$ \\
Linear direct quantity & $d+1$ & $5$ \\
Linear capped direct quantity & $d+1$ & $5$ \\
Linear gated sigmoid direct quantity & $2(d+1)$ & $10$ \\
Linear soft-gated direct quantity & $2(d+1)$ & $10$ \\
Linear hard-gated direct quantity & $2(d+1)$ & $10$ \\
Soft tree, depth-1 linear leaf & $3(d+1)$ & $15$ \\
Soft tree, depth-2 linear leaf & $7(d+1)$ & $35$ \\
\hline
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\small
\caption{Trainable parameter counts for the one-hidden-layer NN scalar-quantity policy families.
These counts use the current repo default $H=50$ for the \texttt{nn\_*} policies. The last column
again reports vanilla / fixed-cost when the decoder-side cap differs across the two canonical
instances.}
\label{tab:nn-parameter-counts}
\begin{tabular}{|l|c|c|}
\hline
Policy family & Trainable parameters & Count at $L=4$ \\
\hline
NN categorical quantity & $H(d+1) + (H+1)(\bar q+1)$ & $1321 / 2851$ \\
NN sigmoid direct quantity & $H(d+1) + (H+1)$ & $301$ \\
NN direct quantity & $H(d+1) + (H+1)$ & $301$ \\
NN capped direct quantity & $H(d+1) + (H+1)$ & $301$ \\
NN gated sigmoid direct quantity & $H(d+1) + 2(H+1)$ & $352$ \\
NN soft-gated direct quantity & $H(d+1) + 2(H+1)$ & $352$ \\
NN hard-gated direct quantity & $H(d+1) + 2(H+1)$ & $352$ \\
\hline
\end{tabular}
\end{table}

\begin{table}[t]
\centering
\small
\caption{Trainable parameter counts for the soft-tree policy families.}
\label{tab:tree-parameter-counts}
\begin{tabular}{|l|c|c|}
\hline
Policy family & Trainable parameters & Count at $L=4$ \\
\hline
Soft tree, depth-1 linear leaf & $3(d+1)$ & $15$ \\
Soft tree, depth-2 linear leaf & $7(d+1)$ & $35$ \\
\hline
\end{tabular}
\end{table}

\section{Numerical experiments}

\subsection{Vanilla lost sales}

Table~\ref{tab:vanilla-l4-demand-table} reports the full $L=4$, $p=4$ vanilla comparison across
all four demand families on the active linear/tree benchmark surface. All heuristic baselines are
shown explicitly rather than being collapsed into a single best-heuristic row.

\begin{table}[t]
\centering
\small
\caption{Vanilla lost-sales mean costs for the single-instance $L=4$, $p=4$ study across all
four demand families under the current $2000$-iteration / population-$64$ CMA-ES protocol. Bold
entries mark the best value in each demand column.}
\label{tab:vanilla-l4-demand-table}
\begin{tabular}{|l||c|c||c|c|}
\hline
 & \multicolumn{2}{c||}{i.i.d. demand} & \multicolumn{2}{c|}{Correlated demand} \\
\hline
Policy / heuristic & Poisson & Geometric & MMPP2+ & MMPP2- \\
\hline
myopic1 & 5.0627 & 11.3176 & 9.8925 & 5.7245 \\
myopic2 & 4.8168 & 10.8025 & 9.5016 & 5.4150 \\
SVBS & 5.8344 & 13.8596 & 12.2518 & 6.8244 \\
\hline
Linear categorical quantity & 4.8752 & 10.8041 & 8.6411 & 5.7431 \\
Linear sigmoid direct quantity & 4.7557 & 10.6606 & 8.6336 & 5.3673 \\
Linear capped direct quantity & 4.7467 & 10.6426 & 8.5899 & 5.3654 \\
Linear gated sigmoid direct quantity & \textbf{4.7451} & 10.6469 & 8.6206 & 5.3680 \\
Linear soft-gated direct quantity & 4.7481 & 10.6416 & 8.5819 & 5.3732 \\
Linear hard-gated direct quantity & 4.7480 & 10.6542 & 8.5799 & 5.3685 \\
Soft tree, depth-1 linear leaf & 4.7481 & \textbf{10.6391} & 8.5893 & --- \\
Soft tree, depth-2 linear leaf & 4.7518 & 10.6453 & 11.9015 & --- \\
\hline
\end{tabular}
\end{table}

The vanilla side is comparatively forgiving:
\begin{itemize}
    \item every non-categorical policy beats the best heuristic in the Poisson and Geometric
    cases,
    \item the correlated-demand winners remain compact dense direct heads rather than the trees,
    \item the one-head sigmoid direct policy is competitive at population $64$,
    \item the currently stored negative-MMPP2 vanilla bundle does not contain the two tree rows,
    so those cells are marked unavailable.
\end{itemize}

So the full vanilla numerical experiment still says the same thing: once we move away from the
flat categorical head, most direct and tree parameterizations land in essentially the same good
basin. Search geometry matters here, but only weakly compared with the fixed-cost case below.

\subsection{Fixed-cost lost sales}

Table~\ref{tab:fixed-cost-l4-demand-table} reports the corresponding fixed-cost $L=4$, $p=4$,
$K=5$ comparison on the same active linear/tree benchmark surface.

\begin{table}[t]
\centering
\small
\caption{Fixed-cost lost-sales mean costs for the single-instance $L=4$, $p=4$, $K=5$ study
across all four demand families under the current $2000$-iteration / population-$64$ CMA-ES
protocol. Bold entries mark the best value in each demand column.}
\label{tab:fixed-cost-l4-demand-table}
\begin{tabular}{|l||c|c||c|c|}
\hline
 & \multicolumn{2}{c||}{i.i.d. demand} & \multicolumn{2}{c|}{Correlated demand} \\
\hline
Policy / heuristic & Poisson & Geometric & MMPP2+ & MMPP2- \\
\hline
$(s,S)$ & 9.3677 & 14.1229 & 12.1579 & 9.8755 \\
$(s,nQ)$ & 9.2094 & 14.0669 & 12.2139 & 9.6925 \\
modified $(s,S,q)$ & 9.2003 & 14.0023 & 12.0935 & 9.6501 \\
\hline
Linear categorical quantity & 10.2754 & 15.9951 & 14.4001 & 10.7435 \\
Linear sigmoid direct quantity & 8.7728 & 15.6229 & 12.3011 & 9.0772 \\
Linear capped direct quantity & 9.7482 & 13.8152 & 12.4235 & 10.3640 \\
Linear gated sigmoid direct quantity & 9.7474 & 13.7711 & 13.2140 & 10.3766 \\
Linear soft-gated direct quantity & 8.7715 & 14.2049 & 12.3907 & \textbf{9.0656} \\
Linear hard-gated direct quantity & 9.7478 & \textbf{13.6276} & \textbf{11.8787} & 10.3620 \\
Soft tree, depth-1 linear leaf & 8.7708 & 13.8218 & 12.3165 & 10.3697 \\
Soft tree, depth-2 linear leaf & \textbf{8.7700} & 13.7412 & 12.3836 & 10.3675 \\
\hline
\end{tabular}
\end{table}

This is still the sharper architecture test:
\begin{itemize}
    \item there is no single winner across all fixed-cost demand families,
    \item flat categorical quantity remains weak in every case,
    \item positively correlated and overdispersed fixed-cost demand favor the hard-gated direct
    head,
    \item negatively correlated fixed-cost demand favors the soft-gated direct head,
    \item Poisson fixed-cost demand still slightly favors the depth-$2$ tree.
\end{itemize}

This is where the full numerical experiment becomes much more informative than a single summary
table. In fixed-cost lost sales, changing the action head changes which basin CMA-ES reaches. The
problem is not just representational capacity: the main empirical effect is that different heads
expose different order/no-order geometry to the search procedure.

\section{Additional single-instance benchmarks}

This section records the first single-instance learned-policy benchmarks for the other Rust-side
problems we have already validated. The goal here is narrower than in the lost-sales study above:
for each problem we state the control model, the learned policy class, and the benchmark outcome on
the current canonical instance.

\subsection{Perishable inventory}

\paragraph{Formulation.}
We use the age-structured single-item perishable inventory model with stochastic demand, positive
lead time, and issuing discipline fixed to FIFO for the benchmark instance. At period $t$, the
state contains the pipeline orders and the on-hand inventory split by remaining lifetime. After the
controller chooses an order quantity $q_t \in \mathbb{Z}_{\ge 0}$, demand is realized, units age,
and expired stock is discarded. The one-period cost is
\[
c q_t + h H_t + p B_t + w W_t,
\]
where $H_t$ is end-of-period holding inventory, $B_t$ is unmet demand, and $W_t$ is waste. The
reported benchmark numbers are mean discounted return, so higher is better.

\paragraph{Policy.}
The learned policy is a depth-$2$ oblique soft decision tree with linear leaves. Each leaf outputs
an affine scalar score in the normalized state features, and the scalar action head uses the
uncapped nonnegative linear decoder already adopted on the Rust side for scalar-action problems.
For this benchmark the trained policy is \texttt{d=2, split=oblique, leaf=linear}.

\paragraph{Results.}
The benchmark instance is \texttt{de\_moor2022\_m4\_exp6\_l2\_cp7\_fifo} with shelf life $4$, lead
time $2$, demand mean $4$, coefficient of variation $0.5$, shortage cost $5$, holding cost $1$,
waste cost $7$, and procurement cost $3$. This problem family is literature-verified: on two
smaller De Moor verification instances, our exact value-iteration return and best base-stock level
match the published tables exactly. On the medium benchmark instance, the trained tree improves on
the two heuristic baselines but still trails the published exact value-iteration benchmark.

\begin{table}[t]
\centering
\caption{Single-instance perishable-inventory benchmark on
\texttt{de\_moor2022\_m4\_exp6\_l2\_cp7\_fifo}. Higher mean discounted return is better.}
\label{tab:perishable-single-instance}
\begin{tabular}{lrr}
\toprule
Policy & Mean discounted return & Gap vs.\ soft tree \\
\midrule
published value iteration & -1432.000 & 14.675 \\
published base stock & -1453.000 & -6.325 \\
repo base stock & -1454.363 & -7.688 \\
repo BSP-low-EW & -1454.363 & -7.688 \\
soft tree, depth-2 linear leaf & \textbf{-1446.675} & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Random-yield inventory}

\paragraph{Formulation.}
We use the single-item random-yield inventory problem with backlogging, positive lead time, and
all-or-nothing delivery success. The state consists of the current inventory level and the pipeline
orders. When the order placed $L$ periods earlier reaches the head of the pipeline, it arrives in
full with probability $\rho$ and otherwise fails completely. After arrival and demand realization,
the one-period cost is
\[
c q_t + h I_t^+ + p (-I_t)^+,
\]
where $I_t$ is the end-of-period inventory position. The reported benchmark numbers are mean
discounted cost, so lower is better.

\paragraph{Policy.}
The learned policy is a depth-$1$ oblique soft decision tree with linear leaves, again using the
uncapped scalar linear decoder on the Rust side. The trained benchmark policy is
\texttt{d=1, split=oblique, leaf=linear}. It is trained with seed-batched rollouts and then
re-evaluated on a disjoint holdout seed set.

\paragraph{Results.}
The benchmark instance is \texttt{yan2026\_style\_lt2\_p075\_discounted} with horizon $12$, lead
time $2$, demand mean $4$, success probability $0.75$, holding cost $1$, shortage cost $9$,
procurement cost $1$, discount factor $0.99$, initial inventory $6$, and initial pipeline
$(4,3)$. This is a repo-native instance shaped after the small-lead-time discounted regime
emphasized by Yan et al., not a verbatim published table row. We therefore do \emph{not} claim
literature verification for this benchmark instance; instead, correctness is anchored by a smaller
exact-DP verifier in the same model family, where the repo exact solution and heuristic costs are
frozen from the solver and match the stored assertions exactly.

\begin{table}[t]
\centering
\caption{Single-instance random-yield benchmark on
\texttt{yan2026\_style\_lt2\_p075\_discounted}. Lower mean discounted cost is better.}
\label{tab:random-yield-single-instance}
\begin{tabular}{lrr}
\toprule
Policy & Mean discounted cost & Gap vs.\ soft tree cost \\
\midrule
linear inflation & 203.762 & 12.022 \\
weighted newsvendor & 222.881 & 31.142 \\
soft tree, depth-1 linear leaf & \textbf{191.740} & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Joint replenishment}

\paragraph{Formulation.}
We use the finite-horizon discounted joint replenishment problem with two items, shared truck
capacity, item-specific minor ordering costs, and holding and shortage penalties. At each period
$t$ the controller selects an order vector $q_t \in \mathbb{Z}_{\ge 0}^n$. If the total ordered
quantity exceeds one truckload, multiple major orders are charged, so the ordering term is
\[
K \left\lceil \frac{\sum_{i=1}^n q_{i,t}}{C} \right\rceil
+ \sum_{i=1}^n k_i \mathbf{1}\{q_{i,t} > 0\}.
\]
After stochastic demand is realized independently for each item, holding and shortage costs are
charged on the ending inventories. The reported benchmark numbers are mean discounted cost, so
lower is better.

\paragraph{Policy.}
The learned policy is a depth-$2$ oblique soft decision tree with linear leaves. Each leaf outputs
an affine vector in the normalized state features, and the resulting action is projected back into
the feasible order box because this problem has vector-valued bounded actions. For this benchmark
the trained policy is \texttt{d=2, split=oblique, leaf=linear}.

\paragraph{Results.}
The current benchmark instance is the repo exact verifier built from the small-scale setting family
used by Vanvuchelen et al.: horizon $4$, discount factor $0.99$, truck capacity $6$, maximum order
quantities $(12,12)$, initial inventory $(2,0)$, major order cost $75$, minor order costs
$(10,10)$, holding costs $(1,1)$, shortage costs $(19,19)$, and independent uniform item demands
on $\{0,\dots,5\}$ and $\{0,\dots,3\}$. The public paper reports the setting family but not a
public exact benchmark row for repo assertions, so this instance is \emph{not} literature-verified;
the optimal DP cost and heuristic benchmarks are instead frozen from our exact solver. On this
instance the trained tree remains above the exact optimum but substantially outperforms the MOQ and
DYN-OUT heuristics.

\begin{table}[t]
\centering
\caption{Single-instance joint-replenishment benchmark on the repo exact verification instance.
Lower mean discounted cost is better.}
\label{tab:joint-single-instance}
\begin{tabular}{lrr}
\toprule
Policy & Mean discounted cost & Gap vs.\ soft tree cost \\
\midrule
optimal DP & 266.386 & -38.122 \\
minimum order quantity & 386.101 & 81.593 \\
dynamic order-up-to & 383.960 & 79.452 \\
soft tree, depth-2 linear leaf & \textbf{304.508} & 0.000 \\
\bottomrule
\end{tabular}
\end{table}

\section{Interpretation}

The original local hypothesis was that \emph{factorization} or \emph{gating} might be the main
winning ingredient. The numerical evidence is now weaker and more precise than that.

So the best current reading is:
\begin{quote}
Good policies are the ones whose parameterization makes the relevant replenishment map easy enough
for CMA-ES to find under the available budget.
\end{quote}

That map need not always be represented in the same way. Vanilla lost sales is comparatively
forgiving, so many non-categorical direct and tree heads realize essentially the same solution.
Fixed-cost lost sales is not: there the action-head geometry and the demand family much more
strongly determine which replenishment boundary CMA-ES can actually find.

\subsection{Dual sourcing}

The dual-sourcing study (Section~\ref{sec:dual-sourcing}) sharpens the same lesson even further:
policy \emph{design} matters more than raw flexibility. A small soft tree whose leaves live in the
capped-dual-index coordinates matches the optimal static heuristic on every Gijs Figure~9 row (and
beats it on two), whereas looser/wider parameterizations and random CMA-ES initialization do
materially worse. The general reading:
\begin{quote}
Small learned policies can match or outperform strong benchmark heuristics, but only when their
action parameterization respects the problem structure. Larger, more flexible trees can perform
materially worse when that structure is too loose.
\end{quote}

\section{Published verification anchor}

Before relying on the fixed-cost simulator and heuristic stack for learned-policy comparisons, we
validate the implementation against the published example used in
\citet{Bijvank2015ParametricPolicies}. Their Table~1 reports a small Poisson-demand instance with:
\[
L=2,\qquad p=14,\qquad K=5,\qquad h=1,\qquad \mu=5.
\]

In the repo, this validation anchor is:
\begin{itemize}
    \item \texttt{bijvank2015\_table1\_l2\_p14\_k5}
\end{itemize}
and is encoded in
\texttt{invman/problems/lost\_sales\_fixed\_order\_cost/reference\_instances.py}.

\begin{table}[t]
\centering
\caption{Published fixed-cost lost-sales validation example from
\citet{Bijvank2015ParametricPolicies} and our reproduced long-run costs for the reported heuristic
parameters.}
\label{tab:fixed-cost-validation}
\begin{tabular}{llll}
\toprule
Policy & Reported parameters & Published cost & Our reproduced cost \\
\midrule
optimal & --- & 11.46 & --- \\
$(s,S)$ & $(17,23)$ & 11.62 & 11.6217 \\
$(s,nQ)$ & $(17,7)$ & 11.56 & 11.5589 \\
modified $(s,S,q)$ & $(17,23,7)$ & 11.50 & 11.5012 \\
\bottomrule
\end{tabular}
\end{table}

The reproduced costs match the published values to within simulation precision. This is the
literature-backed correctness anchor for the fixed-cost environment and heuristic implementation.

It is important to distinguish this validation role from the architecture study above:
\begin{itemize}
    \item the validation instance is used to verify that the simulator and heuristic policies match
    the published benchmark family,
    \item the canonical $L=4$, $p=4$ studies are where we compare modern learned-policy
    parameterizations.
\end{itemize}

\section{Dual sourcing}
\label{sec:dual-sourcing}

\subsection{Problem description}

We study the dual-sourcing inventory model of \citet{gijsbrechts2022drl}: a single item is
replenished from a slow \emph{regular} supplier with lead time $l_r$ and a fast \emph{expedited}
supplier with lead time $l_e=0$, against i.i.d.\ demand. Each period the controller chooses a
regular order $q^r_t$ (unit cost $c_r$) and an expedited order $q^e_t$ (unit cost $c_e>c_r$);
end-of-period net inventory incurs a holding cost $h$ per unit on hand or a backorder penalty $b$
per unit short. Because only the expedite lead time is zero, the relevant post-decision state
reduces to the net inventory plus the outstanding regular-order pipeline
$(I_t, q^r_{t-l_r+1},\dots,q^r_{t-1})$. The classical structured policies are the single-index,
dual-index \citep{veeraraghavan2008dualindex}, capped dual-index, and tailored base-surge
heuristics \citep{sheopuri2010dualsourcing}.

\subsection{Policy functions}

As in the multi-echelon study, the action space is a property of the policy we design. The learned
policy is a soft tree on the reduced pipeline state whose leaves emit a control in the
\emph{factorized capped-dual-index coordinates}
\[
(s_e,\ \Delta_r,\ \bar c_r),\qquad s_r=s_e+\Delta_r,
\]
where $s_e$ is the expedite order-up-to level, $s_r$ the regular order-up-to level, and $\bar c_r$ a
cap on the regular order: the expedited order raises net inventory toward $s_e$ and the regular
order raises the inventory position toward $s_r$ subject to the cap. We compare two geometries: a
\emph{small-cap} variant with $\bar c_r$ on the discrete grid $\{1,2,3,4,6,8,12\}$ (axis-aligned,
constant leaves) and a continuous capped-dual-index-targets variant. This keeps the learned policy
in the heuristic's coordinate system while letting the targets depend on state; raw direct-order
outputs cannot express the capped dual-index control. Because the best \emph{static} control in
these coordinates is the capped dual-index (below), we \emph{warm-start} CMA-ES at the encoded
capped-dual-index solution and let it refine around it --- without the warm start CMA-ES
occasionally lands in a worse basin (the $c_e=110$ rows degrade by $\sim0.5\%$). The wins are thus a
matter of action geometry and reliable optimization, not extra tree capacity.

\subsection{Experiment setup}

We use the six small-scale benchmark instances of \citet{gijsbrechts2022drl} (their Figure~9):
$l_r\in\{2,3,4\}$ and $c_e\in\{105,110\}$, with $l_e=0$, $c_r=100$, $h=5$, $b=495$, demand
$U\{0,1,2,3,4\}$, and order caps $12$ (Table~\ref{tab:ds-params}). The published comparators are the
optimality gaps (relative to the bounded-DP optimum) of the single-index, dual-index, capped
dual-index, and tailored base-surge heuristics and the A3C learner from \citet{gijsbrechts2022drl}:
the capped dual-index gap is $\le0.11\%$ on every row --- it is the best heuristic and an effective
optimal proxy --- whereas A3C ranges from $0.51\%$ to $1.85\%$. We therefore benchmark the learned
policy against the \emph{capped dual-index} cost as the optimal proxy (we do not use our own
bounded DP as the denominator; see Appendix~\ref{app:ds-validation}). Policies are trained with
warm-started CMA-ES, and every policy is evaluated under a common-random-number protocol --- the
same $70$ demand-path seeds and horizon $6\times10^4$ for all policies --- so the dominant per-seed
variance cancels and the paired cost difference is statistically meaningful (paired SEM
$\sim0.002$--$0.006$ versus per-seed standard deviation $\sim0.5$).

\begin{table}[t]
\centering
\caption{Dual-sourcing benchmark instances \citep{gijsbrechts2022drl}. Common to all:
$l_e=0$, $c_r=100$, $h=5$, $b=495$, demand $U\{0,\dots,4\}$, order caps $12$.}
\label{tab:ds-params}
\begin{tabular}{lcc}
\toprule
Instances & $l_r$ & $c_e$ \\
\midrule
dual\_l$\{2,3,4\}$\_ce105 & $2,3,4$ & $105$ \\
dual\_l$\{2,3,4\}$\_ce110 & $2,3,4$ & $110$ \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Results}

Table~\ref{tab:ds-results} compares the best learned policy against the capped dual-index (CDI)
optimal proxy under the common-random-number protocol. The learned soft tree \emph{matches} CDI on
all six instances (four within the discrete-grid rounding floor, $\le+0.003\%$) and \emph{beats} it
on two ($l_r=2,c_e=110$ and $l_r=4,c_e=110$) by a small but statistically significant margin that is
confirmed on held-out seeds. Since CDI is the optimal static policy and the published
$\sim0\%$-gap proxy for the optimum, matching or beating it on every row means the learned policy
operates at the optimum --- and therefore \emph{decisively beats the published A3C baseline} (gaps
$0.51$--$1.85\%$).

\begin{table}[t]
\centering
\caption{Dual-sourcing: best learned soft-tree policy vs.\ the capped dual-index (CDI) optimal
proxy, common-random-number evaluation ($70$ seeds, horizon $6\times10^4$). $\Delta\%$ is the
learned policy's cost relative to CDI (negative is better); the last column is the published A3C
optimality gap from \citet{gijsbrechts2022drl}. The two ``beats'' margins are confirmed on held-out
seeds.}
\label{tab:ds-results}
\begin{tabular}{lrrlr}
\toprule
Instance & CDI (proxy) & Learned & $\Delta\%$ & Pub.\ A3C gap \\
\midrule
$l_r2$ $c_e105$ & 216.806 & 216.806 & $+0.000$ (match) & $0.52\%$ \\
$l_r2$ $c_e110$ & 219.821 & 219.800 & $\mathbf{-0.009}$ (beats) & $0.80\%$ \\
$l_r3$ $c_e105$ & 216.900 & 216.906 & $+0.003$ (match) & $0.82\%$ \\
$l_r3$ $c_e110$ & 220.483 & 220.483 & $+0.000$ (match) & $0.51\%$ \\
$l_r4$ $c_e105$ & 216.914 & 216.918 & $+0.002$ (match) & $1.85\%$ \\
$l_r4$ $c_e110$ & 220.879 & 220.789 & $\mathbf{-0.041}$ (beats) & $1.33\%$ \\
\bottomrule
\end{tabular}
\end{table}

\section{Multi-echelon inventory control: one warehouse, $K$ retailers with special delivery}

\subsection{Problem description}

We study the divergent two-echelon system of \citet{vanroy1997ndp}, also used as a deep-RL
benchmark by \citet{gijsbrechts2022drl}: a single capacitated warehouse replenishes $K$ identical
capacitated retailers. The warehouse orders from an external manufacturer with lead time $l_w$ and
ships to retailers with lead time $l_r$. Each period $t$: (i) the warehouse places an order; (ii)
each retailer raises its inventory position toward a target, shipped from current warehouse stock
(allocated to minimize the spread of retailer positions when warehouse stock is short); (iii)
retailer demand $d_i$ is realized and served from retailer on-hand; (iv) each unmet unit requests a
same-day \emph{special delivery} from leftover warehouse stock with probability $P_w$ and is
otherwise lost --- a hybrid of backlogging and lost sales; (v) pipelines advance. The state is the
warehouse and per-retailer on-hand levels and outstanding-order pipelines. With $I^w_t,I^i_t$ the
end-of-period on-hand inventories and $E_t$ the number of expedited units, the period cost is
\[
c_t = h^w I^w_t + h^r \sum_{i=1}^{K} I^i_t + c^w E_t + p\,(\text{lost units}),
\]
with warehouse/retailer holding costs $h^w,h^r$, unit expedite cost $c^w$, and lost-sale penalty
$p$. Capacities bound the per-period production ($C^m$) and the warehouse and retail inventory
positions ($C^w,C^r$). This is a member of the broader divergent / partial-lost-sales multi-echelon
family \citep{nahmias1994partial, federgruen1984multilocation, dekok2018typology, kaynov2024owmr}.

\subsection{Policy functions}

The action space is a property of the policy we design, not a fixed external grid. We train for the
problem and use the published numbers only as a benchmark; our training is independent of the
learner in \citet{gijsbrechts2022drl}. The policy is a soft decision tree with oblique splits and
linear leaves. It consumes the \emph{pure decision state} (the raw, unnormalized warehouse and
retailer on-hand and pipeline entries) and applies a \emph{policy-owned} divide-by-scale
normalization before acting, mirroring our lost-sales policy interface (the environment emits the
raw state; the policy normalizes it). The two leaf outputs map to a state-dependent warehouse
order-up-to level $y^w_{S_t}$ and a shared retailer order-up-to level $y^r_{S_t}$; the warehouse
order is $q^w_t = \big[\min(y^w_{S_t},C^w) - \mathrm{IP}^w_t\big]^+$ capped by $C^m$, and each
retailer order raises its position toward $\min(y^r_{S_t},C^r)$.

We compare two action designs as a search dimension:
\begin{itemize}
    \item \textbf{grid}: the levels are chosen from the reduced discrete set used by
    \citet{gijsbrechts2022drl}, $y^w\in\{50,60,\dots,100\}$ and $y^r\in\{0,5,\dots,40\}$;
    \item \textbf{direct} (ours): the leaves \emph{directly estimate} the levels as continuous
    values rounded to non-negative integers, bounded only by the physical caps $y^w\in[0,C^w]$ and
    $y^r\in[0,C^r]$.
\end{itemize}
The reduced grid caps the warehouse level at $100$, but the cost-minimizing warehouse base-stock for
these settings is $\approx 300$--$525$ (the published constant base-stock uses $y^w=330,460$); the
grid therefore cannot express the operating region and is a policy artifact, not a problem
constraint.

\subsection{Experiment setup}

We use the two complex case studies of \citet{vanroy1997ndp}, reused as settings~1 and~2 by
\citet{gijsbrechts2022drl} (Table~\ref{tab:me-params}). Both have $K=10$, $h^w=h^r=3$, $c^w=0$,
$p=60$, $P_w=0.8$, $C^m=100$, $C^w=1000$, $C^r=100$; retailer demand is $\mathcal{N}(\mu,\sigma^2)$
rounded to a non-negative integer.

\begin{table}[t]
\centering
\caption{Faithful Gijsbrechts/Van Roy settings (\texttt{gijs\_2022} dynamics).}
\label{tab:me-params}
\begin{tabular}{lcccc}
\toprule
Setting & $l_w$ & $l_r$ & $\mu$ & $\sigma$ \\
\midrule
1 & 2 & 2 & 5 & 14 \\
2 & 5 & 3 & 0 & 20 \\
\bottomrule
\end{tabular}
\end{table}

The published benchmarks we compare against are the constant base-stock costs $(330,23)\to 1302$
and $(460,22)\to 1449$, the best neuro-dynamic-programming costs $1179$ and $1318$, and the A3C
improvements over constant base-stock of $8.95\%\pm0.13\%$ and $12.09\%\pm0.39\%$
\citep{vanroy1997ndp, gijsbrechts2022drl}. As the in-environment benchmark we compute the best
\emph{constant} (state-independent) base-stock by grid search over the physical operating region
(warehouse up to $\approx C^w$, retailer up to $C^r$), \emph{not} the reduced grid. Policies are
trained with CMA-ES; we sweep the action design $\times$ tree depth $\in\{2,3\}$ and evaluate the
long-run average cost over a horizon of $3\times 10^4$ periods and $6$ seeds.

\subsection{Results}

Table~\ref{tab:me-results} reports the best in-environment constant base-stock, the best learned
policy under each action design, and the learned policy's improvement over the constant base-stock,
alongside the published A3C improvement. The direct-estimation policy improves on the best constant
base-stock by $\approx 14.4\%$ on both settings, \emph{exceeding} the published A3C improvements
($8.95\%$ and $12.09\%$). The grid-action policy, restricted to $y^w\le 100$, stays about $230\%$
\emph{above} the benchmark --- the action-space trap, not an environment or learning failure. The
best tree depth is instance-dependent (depth~2 for setting~1, depth~3 for setting~2), which the
design~$\times$~depth sweep selects automatically.

\begin{table}[t]
\centering
\caption{Multi-echelon learned-policy results (\texttt{gijs\_2022} dynamics; mean cost, lower is
better). ``Improvement'' is the learned policy's cost reduction relative to the best constant
base-stock; the last column is the published A3C improvement over constant base-stock from
\citet{gijsbrechts2022drl}.}
\label{tab:me-results}
\begin{tabular}{lccccc}
\toprule
Setting & Best const.\ base-stock $(y^w,y^r)$ & Grid tree & Direct tree (ours) & Improvement & Pub.\ A3C \\
\midrule
1 & $(300,25)\to 911.4$ & $3085.7$ & $\mathbf{779.8}$ & $\mathbf{14.4\%}$ & $8.95\%$ \\
2 & $(525,25)\to 1137.8$ & $3795.1$ & $\mathbf{973.6}$ & $\mathbf{14.4\%}$ & $12.09\%$ \\
\bottomrule
\end{tabular}
\end{table}

\appendix

\section{Environment validation against published costs}

Before using the multi-echelon environment for learned-policy comparison, we validate it against
the published \emph{constant base-stock} costs of \citet{vanroy1997ndp} (the simple problem and the
two complex case studies), evaluated at the published $(y^w,y^r)$ levels under the original
Van-Roy-style cost convention (\texttt{van\_roy\_1997} dynamics), with $100$ replications of horizon
$10^5$, no warm-up discard, and the minimum-spread allocation. Table~\ref{tab:me-validation} shows
all three reproduce within $\sim 1\%$.

\begin{table}[t]
\centering
\caption{Reproduction of the published Van Roy constant base-stock costs.}
\label{tab:me-validation}
\begin{tabular}{lcccc}
\toprule
Instance & Levels $(y^w,y^r)$ & Published & Reproduced & Gap \\
\midrule
Simple problem & $(10,16)$ & $51.7$ & $51.72$ & $+0.03\%$ \\
Case study 1 & $(330,23)$ & $1302$ & $1284.9$ & $-1.31\%$ \\
Case study 2 & $(460,22)$ & $1449$ & $1437.6$ & $-0.79\%$ \\
\bottomrule
\end{tabular}
\end{table}

Two caveats keep this an implementation-fidelity check rather than a full literature verification.
First, exact reproduction requires calibrated demand inputs: the simple problem uses a latent mean
of $6.294$ (the effective mean of Van Roy's $\mathcal{N}(5,8^2)$ after rounding and clipping to
non-negative integers), and case study~2 a latent mean of $1.0$ rather than the stated $0$. Second,
we reproduce only the constant base-stock rows, not the A3C learner. The residual $\sim 1\%$ is
attributed to unspecified details of Van Roy's original simulation protocol. The paper-faithful
settings used for policy design (Table~\ref{tab:me-params}) instead use the \texttt{gijs\_2022} cost
convention and the Table-3 demand means.

\section{Dual-sourcing comparator and the bounded-DP optimum}
\label{app:ds-validation}

For dual sourcing we benchmark against the capped dual-index heuristic as the optimal proxy rather
than a computed optimum. An exhaustive grid search over the capped-dual-index control
$(s_e,\Delta_r,\bar c_r)$ under the evaluation protocol recovers the published capped-dual-index
parameters and confirms it is the best static control on every Figure~9 row, consistent with its
published $\le0.11\%$ optimality gap. We do \emph{not} use our own bounded-DP value iteration as the
denominator: at inventory bounds that keep the $l_r=4$ state space tractable, the truncated
bounded-DP cost is mildly inconsistent (it falls slightly below the simulated heuristic costs), and
widening the bounds to remove the truncation is computationally prohibitive at $l_r=4$. The capped
dual-index proxy is exact to within its published gap and is evaluated on the identical
common-random-number rollout as the learned policies.

\bibliographystyle{plainnat}
\bibliography{references}

\end{document}
